---
title: Ecology of foam establishment during anaerobic digestion of wastewater sludge
  following process disturbance - Supplementary figures
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

## Load packages and data

```{r, message=FALSE, warning=FALSE}
options("lubridate.week.start" = 3)
library(tidyverse)
theme_set(theme_bw())  
library(ggpubr)
library(lubridate)
library(ggrepel)
library(phyloseq)
library(grid)
library(ampvis2)
`%notin%` = Negate(`%in%`)
# Prevalence table function
# get overview of abundances, mean prevalence is the mean 'appearance' of ASVs of the taxon across all samples
prevalencedf <- dget("./custom-functions/myprevalencetablefunction.R")
#pca_helper <- dget("./custom-functions/pca_helper.R")
startoperation <- ymd("2023-03-01")
cols <- c("#4D98AC", "#985C64", "#ff0000", "purple","#C4A484")
names(cols) <- c("Control", "Treatment", "Full-scale", "PSTWAS", "Foam")

## LOAD masterdata
# change this path to where you store the .csv file. you can provide an absolute path like this. Check the path syntax for Windows as it might be different. 
masterdata <- read.csv("./data/masterdataProject1C_Jan25.csv")
masterdata <-  masterdata %>%
  mutate(Date = lubridate::dmy(as.character(Date)) )
# Make AD, Treatment and SludgeType = factors
masterdata$AD <- factor(masterdata$AD, levels = c("AD1", "AD2","AD3", "AD4","AD5", "AD6", "Full-scale", "PSTWAS"))
masterdata$SludgeType <- factor(masterdata$SludgeType, levels = c("Control", "Treatment", "Foam", "Full-scale", "PSTWAS"))
masterdata$Treatment <- factor(masterdata$Treatment, levels = c("Control", "Treatment", "Full-scale", "PSTWAS"))
masterdata$Period <- factor(masterdata$Period, levels = c("Converging", "SteadyState", "Glycerol", "Inhibition", "Recovery/Feedingpause", "Recovery/Foaming", "Recovery/Postfoam", "SteadyState/Postfoam"))
masterdata <- masterdata %>% 
mutate(across(Observed:pielou_ev, ~as.numeric(.)))

# Phyloseq object with 16S abundance data 
ps_1C <- readRDS("./data/ps_215samplesJul24") # Midas53
sample_data(ps_1C)$CSTR <- str_replace_all(sample_data(ps_1C)$AD, c("AD1"="C1", 
                                       "AD2"="T1", 
                                          "AD3"="C2", 
                                            "AD4"="C3", 
                                              "AD5"="T2", 
                                               "AD6"="T3",
                                                "AD_full" = "Full-scale") )
ps.flt  <-  prune_taxa(taxa_sums(ps_1C) >= 10, ps_1C) #minimum reads per feature
prev.genus <- prevalencedf(ps.flt, Genus)

# Phyloseq object with metagenome abundance data 
psmg <- readRDS("./data/ps-metagenome16samples")

## METADATA FOR METAGENOMICS SAMPLES 
# merge individual samples into the pooled samples - mean values
df <- masterdata %>% 
   as_tibble() %>% dplyr::filter(SampleID.DNA %in% c("S1001_AD1", "S1003_AD3", "S1004_AD4",
                                                     "S1002_AD2","S1005_AD5","S1006_AD6",
                                                     "S1147_AD1", "S1149_AD3", "S1150_AD4",
                                                     "S1148_AD2", "S1151_AD5", "S1152_AD6",
                                                     "S1154_AD2","S1157_AD5","S1158_AD6"
                                                     )) %>% column_to_rownames("SampleID.DNA") 
df <- data.frame(df %>% group_by(Date, SludgeType ) %>% summarise_each(funs(mean)))
rownames(df) <- c("S1001_3_4_AD1_3_4", "S1002_5_6_AD2_5_6","S1147_49_50_AD1_3_4", "S1148_51_52_AD2_5_6", "S1154_57_58_AD2_5_6")
df$Period <- c("Inhibition","Inhibition","Recovery/Foaming","Recovery/Foaming", "Recovery/Foaming")
df$AD <- c("AD1.3.4","AD2.5.6","AD1.3.4","AD2.5.6", "AD2.5.6")

# there was also 1 samples pooled for all 6 digesters 
df1 <- masterdata %>% 
   as_tibble() %>% dplyr::filter(SampleID.DNA %in% c("S561_AD1", "S562_AD2", "S563_AD3",
                                                     "S564_AD4","S565_AD5","S566_AD6"
                                                     )) %>% column_to_rownames("SampleID.DNA") 
df1 <- data.frame(df1 %>% group_by(Date ) %>% summarise_each(funs(mean)))
rownames(df1) <- c("S561toS566_AD.AD6")
df1$Period <- c("SteadyState")
df1$AD <- c("AD1.2.3.4.5.6")
# not pooled samples 
df2 <- masterdata %>% 
   as_tibble() %>% dplyr::filter(SampleID.DNA %in% c("F5","F55","F110","F124","F179", 
                                                     "S119_full","S370_full", "S1273_ADFull", "S1554_ADFull", "S1813_ADFull"
                                                     )) %>% 
   column_to_rownames("SampleID.DNA")
#combine the dfs
metadata <- rbind(df,df1, df2)
#rownames(metadata)
metadata <- metadata %>% rownames_to_column("ID") %>% 
  mutate("SampleID" = ID) %>% column_to_rownames("ID")  %>% 
  mutate("SampleID" = factor(SampleID, levels = c("S561toS566_AD.AD6", "S1001_3_4_AD1_3_4", "S1002_5_6_AD2_5_6", "S1147_49_50_AD1_3_4", "S1148_51_52_AD2_5_6", "S1154_57_58_AD2_5_6",   "F5",    "F55", "F110","F124", "F179", "S119_full", "S370_full","S1273_ADFull", "S1554_ADFull","S1813_ADFull")))

metadata$ramaciottieID <- str_replace_all(metadata$SampleID, c( "F5"="Chris_01", "F55"="Chris_02", "F110"="Chris_03","F124" = "Chris_04",
                                      "F179" = "Chris_05", "S119.full" = "Chris_06", "S370.full" = "Chris_07", "S1273_ADFull" = "Chris_08",
                                      "S1554_ADFull" = "Chris_09", "S1813_ADFull" = "Chris_10", "S561toS566_AD.AD6" = "Chris_11", "S1001_3_4_AD1_3_4" = "Chris_12",
                                      "S1002_5_6_AD2_5_6" = "Chris_13", "S1147_49_50_AD1_3_4" = "Chris_14", "S1148_51_52_AD2_5_6" = "Chris_15", "S1154_57_58_AD2_5_6" = "Chris_16"))

metadata <- metadata %>% mutate(VFAs = rowSums(dplyr::select(., contains("acid"))) )
metadata[c("S561toS566_AD.AD6"),c("SludgeType")] <- "Control"
rm(df, df1, df2)


## KEGG Pathways to PHYLOSEQ
## Load KO / KEGG pathway files ps object
ko_to_kegg_reference <- read.csv("./data/ko_to_kegg_reference.csv")[-1]
KO_reference <- read.csv("./data/KO_reference.csv")[-1] 
KO_reference2 <- KO_reference %>% dplyr::select(-KO, -KoDescription) %>% 
                  unique() %>% 
                      separate_wider_delim(Pathway, 
                        delim = ":", 
                        names = c("Pathway","ko"), 
                        too_few = "align_end") %>% 
   filter(Pathway != is.na(Pathway) ) %>% 
   mutate_at("ko", str_replace, "]", "") %>% 
   mutate_at("Pathway", str_replace, "PATH", "")  %>%
   mutate_at("Pathway", str_replace, "\\[", "") 
KO_reference2$Pathway <- str_trim(KO_reference2$Pathway, "right")
KO_reference2 <- KO_reference2 %>% column_to_rownames("ko")

# abundances
KO_abundance <- read.csv("./data/KO.abundances_metagenome.csv")[-1] 
KO_abundance <- KO_abundance %>% column_to_rownames("KO") %>% 
    dplyr::filter(rowSums(.) > 0) %>% 
    rownames_to_column("KO")
kegg_abundance <- read.csv("./data/kegg_abundances_metagenome.csv")[-1]  

# Create the phyloseq object 
ps_kegg <- phyloseq(
  otu_table(kegg_abundance %>% column_to_rownames("ko"), taxa_are_rows = T), 
  sample_data(metadata),
  phyloseq::tax_table(as.matrix(KO_reference2))
)


## PICRUST PATHWAYS TO PHYLOSEQ
# "otu_table"
pws <- data.frame(read_tsv("./data/path_abun_unstrat.tsv", col_names = TRUE)) %>% 
  rename(Id = "pathway") 
# "tax_table"
MC_names <- read_tsv("./data/metacyc_pathway_info_full.tsv") %>% 
  dplyr::select(-ALTERNATIVE) %>%
  dplyr::select(-`SUPER-PATHWAYS`)%>% 
  dplyr::select(-NAMES) %>% 
  dplyr::rename(Pathway2 = "COMMON-NAME") 
# "tax_table new"
MC_names_new <- read_tsv("./data/metacyc_pathway_info_23jan.tsv") 
#Combine
MC_names <- MC_names_new %>% dplyr::full_join(MC_names, by = "Id") %>% 
  filter(Id %in% pws$Id) %>% 
  column_to_rownames("Id")

# metadata 
SplIDs <- rownames(sample_data(readRDS("./data/ps_215samplesJul24"))) %>% str_replace_all(".AD", "_AD")
SplIDs <-  SplIDs %>% str_replace_all(".f", "_f")
metadata.pw <- masterdata %>% 
   as_tibble() %>% dplyr::filter(SampleID.DNA %in% SplIDs) %>% column_to_rownames("SampleID.DNA") 
row.names(metadata.pw) <- row.names(metadata.pw) %>% str_replace_all("_AD", ".AD")
pws <- pws %>% 
  column_to_rownames("Id")

ps_pw <- phyloseq(
  otu_table(pws, taxa_are_rows = T), 
  sample_data(metadata.pw),
  phyloseq::tax_table(MC_names %>% as.matrix())
)


## KRAKEN to PHYLOSEQ
pskr <- readRDS("./data/pskr")


## ONP 16SFL to PHYLOSEQ
meta <- read.csv("./data/0824_sample_sheet16FL.csv") %>% column_to_rownames("alias")
meta$Sample_name <- factor(meta$Sample_name, levels = c( "Feed","Full-scale CSTR","CSTR1","CSTR2","CSTR3","CSTR4","CSTR5","CSTR6"))
meta$SampleType <- factor(meta$SampleType, levels = c( "Control","Treatment","Full-scale","PSTWAS"))
df <- read_tsv("./data/0824_abundance_table_species_16FL.tsv")
taxa <- df %>% 
        separate(tax, sep = ";", into = c("Superkingdom","Kingdom","Phylum","Class","Order","Family","Genus","Species"))%>%   
        dplyr::select(Superkingdom,Phylum,Kingdom,Class,Order,Family,Genus,Species) %>% 
        column_to_rownames("Species")
taxa$Species <- rownames(taxa)
otus <- df[,c(2:9)] 
rownames(otus) <- taxa$Species

# Create the phyloseq object 
ps.onp.ncbi.kraken <- phyloseq(
  otu_table(otus, taxa_are_rows = T), 
  sample_data(meta),
  phyloseq::tax_table(as.matrix(taxa))
)


## FOAM DATA
foamdata <- read.csv("./data/foam_potential_list3Copy_CK.csv")
foamdata <-  foamdata %>%
  mutate(Date = lubridate::dmy(as.character(Date)) )
foamdata <-  foamdata %>% column_to_rownames("X.SludgeID")
```


## Supp. figure S3 gflow
```{r message=FALSE, warning=FALSE}
# Daily from masterfile 
nrow(masterdata %>% 
  dplyr::filter(Gasflow_mL > 20000)) # outlier

generate_vector <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 7
  }
  return(result)
}

generate_vectorday <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 1
  }
  return(result)
}

firstdate <- -2  # selected periods for Supplementary
lastdate <- 166
weeky <- -1
  
pexcl <- masterdata %>% 
  dplyr::filter(Gasflow_mL >= 0 &
                  Gasflow_mL < 20000) %>% 
    dplyr::filter(Date != c("2023-06-08")) %>%   # outlier as the heaters turned off that day
 dplyr::filter(AD != "AD4") %>% #AD4 always behaved differently due to the lack of pre-mixing
  mutate(Gasflow_L = Gasflow_mL / 1000) %>% 
  ggline(x = "daysop",
         y = "Gasflow_L", 
         color = "Treatment",
         legend = "right",
         add = "mean_se",
         add.params = list(width = 0.75)) +
  
 scale_color_manual("CSTR", values = cols) +
  theme_light() +
  
  theme(
        #axis.title.x = element_blank(),
        #axis.text.x = element_text(angle = 0,  hjust=1),
        legend.position = "none",
        legend.background=element_rect(fill = alpha("white", 0.5)) )   + 
  
   scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                      minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) )  +
  
  #  scale_x_date(date_breaks = "1 week", date_labels = "%b %d", 
  #       date_minor_breaks = "1 day",
  #        limits = as.Date(c(firstdate, lastdate)) ) +
  
  annotate("text",x = 2, y = weeky, label = "W01 ", size = 2.5) +   #Week 01
  annotate("text",x = 9, y = weeky, label = "W02 ", size = 2.5) +   #Week 02
   annotate("text",x = 16, y = weeky, label = "W03 ", size = 2.5) +   #Week 03
   annotate("text",x = 23, y = weeky, label = "W04 ", size = 2.5) +   #Week 04
   annotate("text",x = 30, y = weeky, label = "W05 ", size = 2.5) +   #Week 05
   annotate("text",x = 37, y = weeky, label = "W06 ", size = 2.5) +   #Week 06
   annotate("text",x = 44, y = weeky, label = "W07 ", size = 2.5) +   #Week 07
  annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
   annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
   annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
   annotate("text",x = 128, y = weeky, label = "W19 ", size = 2.5) +   #Week 19
   annotate("text",x = 135, y = weeky, label = "W20 ", size = 2.5) +   #Week 20
   annotate("text",x = 142, y = weeky, label = "W21 ", size = 2.5) +   #Week 21
   annotate("text",x = 149, y = weeky, label = "W22 ", size = 2.5) +   #Week 22
   annotate("text",x = 156, y = weeky, label = "W23 ", size = 2.5) +    #Week 23
  annotate("text",x = 163, y = weeky, label = "W24 ", size = 2.5) +    #Week 24
  
  annotate("text",x = 19, y = -.4, label = "Converging", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 0, xend = 31,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 47, y = -.4, label = "Steady state", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 31, xend = 62,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 65, y = -.4, label = "Glycerol   ", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 62, xend = 67,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 74, y = -0.4, label = "   Inhibition", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 67, xend = 82,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 86, y = -0.4, label = "Recovery", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 82, xend = 90,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 94, y = -0.4, label = "Foaming ", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 90, xend = 98,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 101, y = -0.4, label = "    Recovery", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 98, xend = 106,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
     annotate("text",x = 138, y = -.4, label = "Steady state post foam", size = 3) +
   annotate("segment", y = 0, yend = 0, x = 98, xend = 164,
            arrow = arrow(ends = "first", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 85, y = 3.65, label = "Feeding paused", size = 3, angle = 90, alpha = 0.6) +
    annotate("rect", ymin = 0, ymax = 6,  xmin = 82, xmax = 89,
             fill = "blue", alpha = .1) +

    annotate("text",x = 94, y = 7, label = "Foam", size = 3, angle = 90, alpha = 0.6) +
    annotate("rect", ymin = 4.5, ymax = 8, xmin = 92, xmax = 97, 
             fill = "blue", alpha = .1) +
  

  
  ylab(expression(paste("Gasflow (L day"^-1*")"))) +
  xlab("Day of operation")

pincl <- masterdata %>% 
  dplyr::filter(Gasflow_mL >= 0 &
                  Gasflow_mL < 20000) %>% 
    dplyr::filter(Date != c("2023-06-08")) %>%   # outlier as the heaters turned off that day
#  dplyr::filter(AD != "AD4") %>% #AD4 always behaved differently due to the lack of pre-mixing
  mutate(Gasflow_L = Gasflow_mL / 1000) %>% 
  ggline(x = "daysop",
         y = "Gasflow_L", 
         color = "Treatment",
         legend = "right",
         add = "mean_se",
         add.params = list(width = 0.75)) +
  
 scale_color_manual("CSTR", values = cols) +
  theme_light() +
  
  theme(
        axis.title.x = element_blank(),
        #axis.text.x = element_text(angle = 0,  hjust=1),
        legend.position = c(0.90, 0.82),
        legend.background=element_rect(fill = alpha("white", 0.5)) )   + 
  
    scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                      minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) )  +
  
  #  scale_x_date(date_breaks = "1 week", date_labels = "%b %d", 
  #       date_minor_breaks = "1 day",
  #        limits = as.Date(c(firstdate, lastdate)) ) +
  
  annotate("text",x = 2, y = weeky, label = "W01 ", size = 2.5) +   #Week 01
  annotate("text",x = 9, y = weeky, label = "W02 ", size = 2.5) +   #Week 02
   annotate("text",x = 16, y = weeky, label = "W03 ", size = 2.5) +   #Week 03
   annotate("text",x = 23, y = weeky, label = "W04 ", size = 2.5) +   #Week 04
   annotate("text",x = 30, y = weeky, label = "W05 ", size = 2.5) +   #Week 05
   annotate("text",x = 37, y = weeky, label = "W06 ", size = 2.5) +   #Week 06
   annotate("text",x = 44, y = weeky, label = "W07 ", size = 2.5) +   #Week 07
    annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
   annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
   annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
    annotate("text",x = 128, y = weeky, label = "W19 ", size = 2.5) +   #Week 19
   annotate("text",x = 135, y = weeky, label = "W20 ", size = 2.5) +   #Week 20
   annotate("text",x = 142, y = weeky, label = "W21 ", size = 2.5) +   #Week 21
   annotate("text",x = 149, y = weeky, label = "W22 ", size = 2.5) +   #Week 22
   annotate("text",x = 156, y = weeky, label = "W23 ", size = 2.5) +    #Week 23
  annotate("text",x = 163, y = weeky, label = "W24 ", size = 2.5) +    #Week 24
  
   annotate("text",x = 19, y = -.4, label = "Converging", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 0, xend = 31,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 47, y = -.4, label = "Steady state", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 31, xend = 62,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 65, y = -.4, label = "Glycerol   ", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 62, xend = 67,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 74, y = -0.4, label = "   Inhibition", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 67, xend = 82,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 86, y = -0.4, label = "Recovery", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 82, xend = 90,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 94, y = -0.4, label = "Foaming ", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 90, xend = 98,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 101, y = -0.4, label = "    Recovery", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 98, xend = 106,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 138, y = -.4, label = "Steady state post foam", size = 3) +
   annotate("segment", y = 0, yend = 0, x = 98, xend = 164,
            arrow = arrow(ends = "first", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 85, y = 3.65, label = "Feeding paused", size = 3, angle = 90, alpha = 0.6) +
    annotate("rect", ymin = 0, ymax = 6,  xmin = 82, xmax = 89,
             fill = "blue", alpha = .1) +

    annotate("text",x = 94, y = 7, label = "Foam", size = 3, angle = 90, alpha = 0.6) +
    annotate("rect", ymin = 4.5, ymax = 8, xmin = 92, xmax = 97, 
             fill = "blue", alpha = .1) +
  
  ylab(expression(paste("Gasflow (L day"^-1*")"))) 

ggarrange(pincl, pexcl, labels = "AUTO", ncol = 1)

ggsave("./Figures/Sup_lineplot_gasdaily_inclexclAD4.png", height=20, width=27, units='cm', dpi=300)
```

## Supp. figure S3 gflow - ind ADs (not used)
```{r fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
# Daily from masterfile 
nrow(masterdata %>% 
  dplyr::filter(Gasflow_mL > 20000)) # outlier

firstdate <- "2023-03-01"  
lastdate <- "2023-08-15"

masterdata %>% 
  dplyr::filter(Gasflow_mL >= 0 &
                  Gasflow_mL < 20000) %>% 
    dplyr::filter(Date != c("2023-06-08")) %>%   # outlier as the heaters turned off that day
#  dplyr::filter(AD != "AD4") %>% #AD4 always behaved differently due to the lack of pre-mixing
  mutate(Gasflow_L = Gasflow_mL / 1000) %>% 
  ggline(x = "Date",
         y = "Gasflow_L", 
         color = "AD",
         legend = "right",
         add = "mean_se",
         add.params = list(width = 0.75)) +
 #scale_color_manual("Sludge Type", values = cols) +
  theme_light() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45,  hjust=1),
        legend.position = c(0.90, 0.82),
        legend.background=element_rect(fill = alpha("white", 0.5)) )   + 
  
    scale_x_date(date_breaks = "1 week", date_labels = "%b %d", 
         date_minor_breaks = "1 day",
          limits = as.Date(c(firstdate, lastdate)) ) +

  annotate("text",x = as.Date("2023-04-21"), y = -1, label = "W08 ", size = 2.5) +   #Week 08
   annotate("text",x = as.Date("2023-04-28"), y = -1, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = as.Date("2023-05-05"), y = -1, label = "W10 ", size = 2.5) +   #Week 10
   annotate("text",x = as.Date("2023-05-12"), y = -1, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = as.Date("2023-05-19"), y = -1, label = "W12 ", size = 2.5) +   #Week 12
   annotate("text",x = as.Date("2023-05-26"), y = -1, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = as.Date("2023-06-02"), y = -1, label = "W14 ", size = 2.5) +   #Week 14
   annotate("text",x = as.Date("2023-06-09"), y = -1, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = as.Date("2023-06-16"), y = -1, label = "W16 ", size = 2.5) +   #Week 16
   annotate("text",x = as.Date("2023-06-23"), y = -1, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = as.Date("2023-06-30"), y = -1, label = "W18 ", size = 2.5) +    #Week 18
  
   annotate("text",x = as.Date("2023-04-24"), y = -.4, label = "Steady state", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = as.Date("2023-04-17"), xend = as.Date("2023-05-02"),
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = as.Date("2023-05-05"), y = -.4, label = "Glycerol   ", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = as.Date("2023-05-02"), xend = as.Date("2023-05-07"),
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = as.Date("2023-05-14"), y = -0.4, label = "   Inhibition", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = as.Date("2023-05-07"), xend = as.Date("2023-05-22"),
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = as.Date("2023-05-26"), y = -0.4, label = "Recovery", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = as.Date("2023-05-22"), xend = as.Date("2023-05-30"),
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = as.Date("2023-06-03"), y = -0.4, label = "Foaming ", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = as.Date("2023-05-30"), xend = as.Date("2023-06-07"),
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = as.Date("2023-06-10"), y = -0.4, label = "    Recovery", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = as.Date("2023-06-07"), xend = as.Date("2023-06-15"),
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  
    annotate("text",x = as.Date("2023-05-25"), y = 4, label = "Paused feeding", size = 2.5, angle = 90, alpha = 0.6) +
    annotate("rect", ymin = 0, ymax = 6,  xmin = as.Date("2023-05-22"), xmax = as.Date("2023-05-29"),
             fill = "blue", alpha = .1) +

    annotate("text",x = as.Date("2023-06-03"), y = 7, label = "Foam", size = 2.5, angle = 90, alpha = 0.6) +
    annotate("rect", ymin = 4.5, ymax = 8, xmin = as.Date("2023-06-01"), xmax = as.Date("2023-06-06"), 
             fill = "blue", alpha = .1) +
  
   annotate("text",x = as.Date("2023-06-24"), y = -.4, label = "Steady state post foam", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = as.Date("2023-06-07"), xend = as.Date("2023-07-03"),
            arrow = arrow(ends = "first", angle = 90, length = unit(.1,"cm"))) +
  
  ylab(expression(paste("Gasflow (L day"^-1*")"))) 

ggsave("./Figures/Sup_lineplot_gasdaily_indADs.png", height=15, width=25, units='cm', dpi=300)
```


## Supp. figure S4 CH4
```{r message=FALSE, warning=FALSE}

firstdate <- -2  # selected periods for Supplementary
lastdate <- 166
weeky <- c(-6.5)
labely <- c(-2.5)
segmenty <- c(0)


generate_vector <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 7
  }
  return(result)
}

generate_vectorday <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 1
  }
  return(result)
}


## Gas comp
masterdata %>%
    dplyr::filter(CH4_pct >= 0) %>% 
  ggline(x = "daysop",
         y = "CH4_pct", 
         color = "Treatment",
         legend = "right",
         add = "mean_se",
         add.params = list(width = 1)) +
  ylab("CH4 (%)") +
  theme_light() +
  scale_color_manual("CSTR", values = cols)  +
  theme(
       #axis.title.x = element_blank(),
       # axis.text.x = element_text(angle = 0,  hjust=1),
        legend.position = c(0.9, 0.4)
       # legend.position = "none" 
        )   + 
  
   scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                      minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) )  +

  annotate("text",x = 2, y = weeky, label = "W01 ", size = 2.5) +   #Week 01
  annotate("text",x = 9, y = weeky, label = "W02 ", size = 2.5) +   #Week 02
   annotate("text",x = 16, y = weeky, label = "W03 ", size = 2.5) +   #Week 03
   annotate("text",x = 23, y = weeky, label = "W04 ", size = 2.5) +   #Week 04
   annotate("text",x = 30, y = weeky, label = "W05 ", size = 2.5) +   #Week 05
   annotate("text",x = 37, y = weeky, label = "W06 ", size = 2.5) +   #Week 06
   annotate("text",x = 44, y = weeky, label = "W07 ", size = 2.5) +   #Week 07
    annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
   annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
   annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
    annotate("text",x = 128, y = weeky, label = "W19 ", size = 2.5) +   #Week 19
   annotate("text",x = 135, y = weeky, label = "W20 ", size = 2.5) +   #Week 20
   annotate("text",x = 142, y = weeky, label = "W21 ", size = 2.5) +   #Week 21
   annotate("text",x = 149, y = weeky, label = "W22 ", size = 2.5) +   #Week 22
   annotate("text",x = 156, y = weeky, label = "W23 ", size = 2.5) +    #Week 23
  annotate("text",x = 163, y = weeky, label = "W24 ", size = 2.5) +    #Week 24
  
   annotate("text",x = 19, y = labely, label = "Converging", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 0, xend = 31,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 47, y = labely, label = "Steady state", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 31, xend = 62,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 65, y = labely, label = "Glycerol   ", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 62, xend = 67,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 74, y = labely, label = "   Inhibition", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 67, xend = 82,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 86, y = labely, label = "Recovery", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 82, xend = 90,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 94, y = labely, label = "Foaming ", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 90, xend = 98,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 101, y = labely, label = "    Recovery", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 98, xend = 106,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
     annotate("text",x = 138, y = labely, label = "Steady state post foam", size = 2.5) +
   annotate("segment", y = 0, yend = 0, x = 98, xend = 164,
            arrow = arrow(ends = "first", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 85.5, y = 35, label = "Feeding\npaused", size = 3.5, alpha = 0.4, angle = 90) +
    annotate("rect", ymin = 0, ymax = 75,  xmin = 82, xmax = 89,
             fill = "blue", alpha = .1) +

    annotate("text",x = 94.5, y = 67, label = "Foaming", size = 2.5, alpha = 0.5, angle = 90) +
    annotate("rect", ymin = 60, ymax = 75, xmin = 92, xmax = 97, fill = "blue", alpha = .1) +
 # 
  ylab(expression(paste("CH"[4]*" (%)"))) +
  xlab("Day of operation")
  
ggsave("./Figures/Sup_lineplot_CH4.png", height=8, width=26, units='cm', dpi=300)
```

## Supp. figure S5 pH
```{r message=FALSE, warning=FALSE}
firstdate <- -2  # selected periods for Supplementary
lastdate <- 166
weeky <- c(4.8)
labely <- c(5.25)
segmenty <- c(5)

generate_vector <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 7
  }
  return(result)
}

generate_vectorday <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 1
  }
  return(result)
}

masterdata %>% 
  dplyr::filter(Treatment != "PSTWAS") %>% 
  dplyr::filter(pH >= 0) %>%  
  ggline(x = "daysop",
         y = "pH", 
         color = "Treatment",
         legend = "right",
         add = "mean_se",
         add.params = list(width = 1)) +
 scale_color_manual("CSTR", values = cols) +
  theme_light() +
  theme(
        #axis.title.x = element_blank(),
        #axis.text.x = element_text(angle = 0,  hjust=1),
        legend.position = c(0.9, 0.4))  + 
  
 scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                      minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) )  +
    
   annotate("text",x = 2, y = weeky, label = "W01 ", size = 2.5) +   #Week 01
   annotate("text",x = 9, y = weeky, label = "W02 ", size = 2.5) +   #Week 02
   annotate("text",x = 16, y = weeky, label = "W03 ", size = 2.5) +   #Week 03
   annotate("text",x = 23, y = weeky, label = "W04 ", size = 2.5) +   #Week 04
   annotate("text",x = 30, y = weeky, label = "W05 ", size = 2.5) +   #Week 05
   annotate("text",x = 37, y = weeky, label = "W06 ", size = 2.5) +   #Week 06
   annotate("text",x = 44, y = weeky, label = "W07 ", size = 2.5) +   #Week 07
    annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
   annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
   annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
    annotate("text",x = 128, y = weeky, label = "W19 ", size = 2.5) +   #Week 19
   annotate("text",x = 135, y = weeky, label = "W20 ", size = 2.5) +   #Week 20
   annotate("text",x = 142, y = weeky, label = "W21 ", size = 2.5) +   #Week 21
   annotate("text",x = 149, y = weeky, label = "W22 ", size = 2.5) +   #Week 22
   annotate("text",x = 156, y = weeky, label = "W23 ", size = 2.5) +    #Week 23
  annotate("text",x = 163, y = weeky, label = "W24 ", size = 2.5) +    #Week 24
  
  annotate("text",x = 19, y = labely, label = "Converging", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 0, xend = 31,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 47, y = labely, label = "Steady state", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 31, xend = 62,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 65, y = labely, label = "Glycerol   ", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 62, xend = 67,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 74, y = labely, label = "   Inhibition", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 67, xend = 82,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 86, y = labely, label = "Recovery", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 82, xend = 90,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 94, y = labely, label = "Foaming ", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 90, xend = 98,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 101, y = labely, label = "    Recovery", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 98, xend = 106,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
     annotate("text",x = 138, y = labely, label = "Steady state post foam", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 98, xend = 164,
            arrow = arrow(ends = "first", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 85.5, y = 7, label = "Feeding\npaused", size = 3.5, alpha = 0.4, angle = 90) +
    annotate("rect", ymin = 6, ymax = 7.75,  xmin = 82, xmax = 89,
             fill = "blue", alpha = .1) +
  
  ylab("pH") +
  xlab("Day of operation")

ggsave("./Figures/Sup_lineplot_ph.png", height=8, width=26, units='cm', dpi=300)

```

## Supp. figure S6 DNA (mg gCOD)
```{r message=FALSE, warning=FALSE}
firstdate <- 47  
lastdate <- 124
weeky <- -0.3
labely <- c(-0.125)
segmenty <- c(0)
  
masterdata %>% dplyr::filter(DNA_mg_gCOD > 0 & 
                               DNA_mg_gCOD < 5 ) %>% 
  ggline(x = "daysop",
         y = "DNA_mg_gCOD", 
         color = "SludgeType",
         legend = "right",
         add = "mean_se",
         add.params = list(width = 1)) +
 
      scale_color_manual("Sludge type", 
        values = cols, labels=c('Control CSTR', 'Treatment CSTR', 'Full-scale CSTR', "PS/TWAS")) +
    scale_shape_manual(values = c(15, 15, 16, 17)) +
  
  guides(color = guide_legend( 
    override.aes=list(shape = c(15, 15, 16, 17))),
    shape = "none") +

  theme_light() +
  
  theme(
      #axis.title.x = element_blank(),
      #axis.text.x = element_text(angle = 0,  hjust=.5)
      legend.position = c(0.88, 0.83),
      legend.background=element_rect(fill = alpha("white", 0.1)),
      legend.key.height = unit(0.7, "line")  # Adjust this value to make the legend shorter
      )  + 
  
     scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                      minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) )  +
  
  annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
   annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
   annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
  
  
 annotate("text",x = 54, y = labely, label = "Steady state", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 47, xend = 62,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 65, y = labely, label = "Glycerol   ", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 62, xend = 67,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 74, y = labely, label = "   Inhibition", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 67, xend = 82,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 85, y = labely, label = "  Recovery", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 82, xend = 90,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 94, y = labely, label = "Foaming ", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 90, xend = 98,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 101, y = labely, label = "  Recovery", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 98, xend = 106,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 115, y = labely, label = "Steady state post foam", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 98, xend = 124,
            arrow = arrow(ends = "first", angle = 90, length = unit(.1,"cm"))) +
  

    annotate("text",x = 85, y = 1.25, label = "Feeding paused", size = 3, angle = 90, alpha = 0.5) +
    annotate("rect", ymin = 0, ymax = 2.5, xmin = 82, xmax = 89, fill = "blue", alpha = .1) +
  
  ylab(expression(paste("DNA (mg gCOD"^-1*")"))) +
  xlab("Day of operation")

ggsave("./Figures/Sup_lineplot_mgCOD.png", height=3, width=6.75, units='in', dpi=300)
```

## Supp. figure S7 Shannon
```{r message=FALSE, warning=FALSE}
firstdate <- 47  
lastdate <- 124
weeky <- 3.2
labely <- c(3.35)
segmenty <- c(3.5)

masterdata %>% dplyr::filter(ngmL_qb >= 0 ) %>% 
  ggline(x = "daysop",
         y = "Shannon", 
         color = "SludgeType",
         legend = "top",
         add = "mean_se",
         add.params = list(width = 1)) +

        scale_color_manual("Sample type", 
        values = cols, labels=c('Control CSTR', 'Treatment CSTR', 'Foam', "Full-scale CSTR", "PS/TWAS")) +
    scale_shape_manual(values = c(15, 15, 16, 15, 17)) +
  
  guides(color = guide_legend( 
    override.aes=list(shape = c(15, 15, 16,  15, 17))),
    shape = "none") +
  
  theme_light() +
 
  theme(
      #axis.title.x = element_blank(),
      #axis.text.x = element_text(angle = 0,  hjust=.5)
   #   legend.position = c(0.88, 0.83),
   #   legend.background=element_rect(fill = alpha("white", 0.1)),
     legend.key.height = unit(0.7, "line")  # Adjust this value to make the legend shorter
     )  + 
  
   scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                      minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) )  +

    annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
   annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
   annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
  
  
   annotate("text",x = 54, y = labely, label = "Steady state", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 47, xend = 62,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 65, y = labely, label = "Glycerol   ", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 62, xend = 67,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 74, y = labely, label = "   Inhibition", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 67, xend = 82,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 85, y = labely, label = "  Recovery", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 82, xend = 90,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 94, y = labely, label = "Foaming ", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 90, xend = 98,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 101, y = labely, label = "  Recovery", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 98, xend = 106,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 115, y = labely, label = "Steady state post foam", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 98, xend = 124,
            arrow = arrow(ends = "first", angle = 90, length = unit(.1,"cm"))) +
  

    annotate("text",x = 85, y = 4.25, label = "Feeding paused", size = 3, angle = 90, alpha = 0.5) +
    annotate("rect", ymin = segmenty, ymax = 5, xmin = 82, xmax = 89, fill = "blue", alpha = .1) +

    #annotate("text",x = as.Date("2023-06-03"), y = 301, label = "Foam", size = 2) +
  #  annotate("rect", ymin = 5.5, ymax = 125, xmin = as.Date("2023-06-01"), xmax = as.Date("2023-06-06"), 
  #           fill = "blue", alpha = .1) +
  ylab(expression(paste("Shannon")))  +
  xlab("Day of operation")

ggsave("./Figures/Sup_lineplot_shannon.png", height=3, width=7.5, units='in', dpi=300)
```


## Supplementary Figure S8
### VFAs
```{r message=FALSE, warning=FALSE}
weeky <- c(-150)
labely <- c(50)
segmenty <- c(0)
generate_vector <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 7
  }
  return(result)
}

generate_vectorday <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 1
  }
  return(result)
}

VFAs <- masterdata %>% 
        dplyr::select(SampleID.VFA, daysop, Date, AD, Treatment, Ethanol, Propanol,Butanol, X1.Hexanol, Acetic.acid, Propionic.acid, iso.Butyric.acid, Butyric.acid, iso.Valeric.acid, Valeric.acid, X4.Methyl.valeric.acid, Hexanoic.acid) %>% filter(SampleID.VFA != "")

# Control
p1 <- VFAs %>% 
  pivot_longer(Acetic.acid:Hexanoic.acid, names_to = "Compound") %>% 
 dplyr::filter(Treatment %in% c("Control")) %>% 
 # dplyr::filter(AD != c("AD3")) %>% 
  group_by(daysop, Compound, Treatment) %>% 
  summarise(value = median(value)) %>% 
ggplot(., aes(x=daysop, y=value, fill=Compound)) + 
    geom_area(alpha=0.6 , linewidth=.1, colour="white", position = "stack") +
  facet_wrap(vars(Treatment)) +
  ylab(expression(paste("Concentration (mg L"^-1*")"))) + 
  scale_fill_viridis_d(labels = c("Acetic acid"," Butyric acid","Hexanoic acid","iso-Butyric acid","iso-Valeric acic","Propionic acid", "Valeric acid", "4-Methylvaleric acid")) +
  theme_light() +
  theme( 
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 0,  hjust=1, vjust = 0.5)
        ) + 
  
   scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                      minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) )  +
  
  annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
   annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
   annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
  
  ylim(weeky, 5000)  +
  
  guides(fill = guide_legend(override.aes = list(size = 4), 
          keywidth = 0.5, 
          keyheight = 0.5))   + 
  
  annotate("text",x = 85, y = 3000, label = "Feeding paused", size = 3, angle = 90, alpha = .6) +
    annotate("rect", ymin = 0, ymax = 5000,  xmin = 82, xmax = 89, fill = "blue", alpha = .1) 

# Treatment
p2 <- VFAs %>% pivot_longer(Acetic.acid:Hexanoic.acid, names_to = "Compound") %>% 
 dplyr::filter(Treatment %in% c("Treatment")) %>% 
  group_by(daysop, Compound, Treatment) %>% 
  summarise(value = median(value)) %>% 
ggplot(., aes(x=daysop, y=value, fill=Compound)) + 
    geom_area(alpha=0.6 , linewidth=.1, colour="white", position = "stack") +
  facet_wrap(vars(Treatment)) +
  #ylab("Concentration (mg/L)") +
  scale_fill_viridis_d(labels = c("Acetic acid"," Butyric acid","Hexanoic acid","iso-Butyric acid","iso-Valeric acic","Propionic acid", "Valeric acid", "4-Methylvaleric acid")) +
  theme_light() +
  theme(
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 0,  hjust=1, vjust = 0.5))  + 
  
   scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                      minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) )  +
  
  annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
   annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
   annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
  
    annotate("text",x = 64, y = 3000, label = "Glycerol added", size = 3, angle = 90, alpha = .6) +
    annotate("rect",ymin =0, ymax = 5000, xmin = 62, xmax = 66, fill = "blue", alpha = .1) +

    annotate("text",x = 85, y = 3000, label = "Feeding paused", size = 3, angle = 90, alpha = .6) +
    annotate("rect", ymin = 0, ymax = 5000,  xmin = 82, xmax = 89, fill = "blue", alpha = .1) +

    annotate("text",x = 94, y = 3000, label = "Foam observed", size = 3, angle = 90, alpha = .6) +
    annotate("rect", ymin = 0, ymax = 5000, xmin =92, xmax = 97, fill = "blue", alpha = .1)  +
  
  guides(fill = guide_legend(override.aes = list(size = 4), 
          keywidth = 0.5, 
          keyheight = 0.5))

g1 <- ggarrange(p1, theme_void(), p2, common.legend = TRUE, nrow = 1, legend = "top", align = "v", 
                widths = c(1,-0.02,0.9) )

g1 <- annotate_figure(g1, bottom = textGrob("Day of operation", gp = gpar(cex = 0.9)))

ggsave(plot = g1,  "./Figures/areaplot_vfas_daysop.pdf", height=9, width=17, units='cm', dpi=300)
```


## Supp. figure S9 CN
```{r message=FALSE, warning=FALSE}
### CN
masterdata %>% 
   dplyr::filter(AD != "PSTWAS" &
                   AD != "Full-scale" &
                   week %in% c(9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24)) %>% 
  # drop_na(CN) %>% 
  #dplyr::filter(CN >= 0) %>% 
  ggboxplot(x = "week",
         y = "CN", 
         fill = "Treatment",
         legend = "right") +
  ylab("C/N ratio") +
  theme_light() +
  scale_fill_manual("Treatment", values = cols)  +
  labs(x = "Week") +
    geom_point(aes(x =week,y=CN, fill=Treatment), 
             position = position_jitterdodge(), alpha = 0.3) +
  scale_x_discrete(breaks = c(9,10,11,12,13,14,15,16,17,18,21,24),
                   labels = c("W09","W10","W11","W12","W13","W14","W15","W16","W17","W18","W21","W24")) +
   annotate("text",x = 5, y = 6, label = "Feeding\npaused", size = 2.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave( "./Figures/Sup_boxplot_CN.png", height=7, width=17,units='cm', dpi=600)

str(masterdata)
```



## Supp. figure S10 foam heights
```{r}
firstdate <- 47  # selected periods for general manuscript
lastdate <- 124

weeky <- c(-30)
labely <- c(-10)
segmenty <- c(0)

generate_vector <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 7
  }
  return(result)
}

generate_vectorday <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 1
  }
  return(result)
}


foamdata %>% 
  dplyr::filter(SludgeType %notin% c("Full-scale")) %>%   # filter columns out
  ggline(x = "DayofOp",
         y = "Max", 
         color = "SludgeType",
         legend = "right",
         add = "mean_se",
         add.params = list(width = 1)) +
  # scale_color_manual(values = cols) +   # create a cols vector for custom colours
  theme_light() +
  scale_color_manual("Sludge Supernatent", values = cols)  +

    theme(
        axis.text.x = element_text(angle = 0,  hjust=0.5),
        legend.position = c(0.90, 0.825),
        legend.background=element_rect(fill = alpha("white", 0.5)),
        legend.ticks.length = element_blank(),
        legend.title = element_blank() )   + 
  
  scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                     minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) )  +

  annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
   annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
   annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
  
  
  annotate("text",x = 54, y = labely, label = "Steady state", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 47, xend = 62,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
 #  annotate("text",x = 65, y = labely, label = "Glycerol   ", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 62, xend = 67,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 74, y = labely, label = "   Inhibition", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 67, xend = 82,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 86, y = labely, label = "Recovery", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 82, xend = 90,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  #annotate("text",x = 94, y = labely, label = "Foaming ", size = 2.5) +
  annotate("segment", y = segmenty, yend = segmenty, x = 90, xend = 98,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 101, y = labely, label = "    Recovery", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 98, xend = 106,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 115, y = labely, label = "Steady state post foam", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 98, xend = 124,
            arrow = arrow(ends = "first", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 85, y = 100, label = "Feeding paused", size = 4, angle = 90, alpha = 0.5) +
    annotate("rect", ymin = 0, ymax = 200, xmin = 82, xmax = 89, fill = "blue", alpha = .1) +

      annotate("text",x = 64, y = 75, label = "Glycerol", size = 4, angle = 90, , alpha = 0.5) +
    annotate("rect",ymin =0.5, ymax = 150, xmin = 62, xmax = 67, 
             fill = "blue", alpha = .1) +

    annotate("text",x = 93.5, y = 75, label = "Foam observed", size = 3.5, angle = 90, , alpha = 0.5) +
    annotate("rect", ymin = 0.5, ymax = 150, xmin = 91, xmax = 96, 
             fill = "blue", alpha = .1) +
  
      ylab("Max foam height (mm)") +
      xlab("Day of operation")

ggsave("./Figures/Sup_lineplot_foam_daysop.png", height=2.5, width=6, units='in', dpi=300)
```

## Supp. figure S12 phylum abund
#### Control
```{r message=FALSE, warning=FALSE}

firstdate <- -2  # selected periods for Supplementary
lastdate <- 166

weeky <- c(-5)
labely <- c(.05)
segmenty <- c(0)

generate_vector <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 7
  }
  return(result)
}

generate_vectorday <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 1
  }
  return(result)
}

ps.rel <- microbiome::aggregate_taxa(ps.flt, "Phylum")

ps.rel <- ps.rel %>% 
  microbiome::transform(transform = 'compositional')
df <- phyloseq::psmelt(ps.rel) %>% 
  mutate(Abundance = Abundance *100)

p1 <- df %>% 
  group_by(daysop, Kingdom, Phylum, CSTR, SludgeType) %>% 
   summarise(value = median(Abundance)) %>% 
    dplyr::filter(SludgeType %in% c("Control")) %>% 
  
  ggplot(., aes(x=daysop, y=value, fill=Phylum)) + 
  geom_area(alpha=0.6 , linewidth=.1, colour="white", position = "stack") +
  facet_wrap(vars(CSTR), ncol = 1) +
  
  scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                      minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) ) +
  
  ylab("Relative Abundance") +
  theme_light() +
 # scale_color_manual("Treatment", values = cols)  +
  theme(
        axis.title.x = element_blank(),
       #axis.text.x = element_blank(),
        axis.text.x = element_text(angle = 90,  hjust=1),
        legend.position = "top",
       legend.title = element_blank()
       )   +

  #  annotate("text",x = 2, y = weeky, label = "W01 ", size = 2.5) +   #Week 01
   annotate("text",x = 9, y = weeky, label = "W02 ", size = 2.5) +   #Week 02
#   annotate("text",x = 16, y = weeky, label = "W03 ", size = 2.5) +   #Week 03
   annotate("text",x = 23, y = weeky, label = "W04 ", size = 2.5) +   #Week 04
#   annotate("text",x = 30, y = weeky, label = "W05 ", size = 2.5) +   #Week 05
   annotate("text",x = 37, y = weeky, label = "W06 ", size = 2.5) +   #Week 06
#   annotate("text",x = 44, y = weeky, label = "W07 ", size = 2.5) +   #Week 07
    annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
#   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
#   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
#   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
#   annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
 #  annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
#    annotate("text",x = 128, y = weeky, label = "W19 ", size = 2.5) +   #Week 19
   annotate("text",x = 135, y = weeky, label = "W20 ", size = 2.5) +   #Week 20
#   annotate("text",x = 142, y = weeky, label = "W21 ", size = 2.5) +   #Week 21
   annotate("text",x = 149, y = weeky, label = "W22 ", size = 2.5) +   #Week 22
#   annotate("text",x = 156, y = weeky, label = "W23 ", size = 2.5) +    #Week 23
  annotate("text",x = 163, y = weeky, label = "W24 ", size = 2.5) +    #Week 24
  
   annotate("text",x = 85, y = 50, label = "Feeding paused", size = 4, angle = 90, alpha = 0.6) +
    annotate("rect", ymin = 0, ymax = 100,  xmin = 82, xmax = 89,
             fill = "blue", alpha = .1) +
  guides(fill=guide_legend(ncol=6)) + 
  ggtitle("Control")

```

#### Treatment
```{r message=FALSE, warning=FALSE}

ps.rel <- microbiome::aggregate_taxa(ps.flt, "Phylum")

ps.rel <- ps.rel %>% 
  microbiome::transform(transform = 'compositional')
df <- phyloseq::psmelt(ps.rel)  %>% 
  mutate(Abundance = Abundance *100)

p2 <- df %>% 
  group_by(daysop, Kingdom, Phylum, CSTR, SludgeType) %>% 
   summarise(value = median(Abundance)) %>% 
  dplyr::filter(SludgeType %in% c("Treatment")) %>% 
  
  ggplot(., aes(x=daysop, y=value, fill=Phylum)) + 
  geom_area(alpha=0.6 , linewidth=.1, colour="white", position = "stack") +
  facet_wrap(vars(CSTR), ncol = 1) +
  
  scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                      minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) ) +
  
  ylab("Relative Abundance") +
  theme_light() +
   theme(
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
         axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90,  hjust=1),
        legend.position = "top",
        legend.title = element_blank()
        )   +
 
 #   annotate("text",x = 2, y = weeky, label = "W01 ", size = 2.5) +   #Week 01
   annotate("text",x = 9, y = weeky, label = "W02 ", size = 2.5) +   #Week 02
#   annotate("text",x = 16, y = weeky, label = "W03 ", size = 2.5) +   #Week 03
   annotate("text",x = 23, y = weeky, label = "W04 ", size = 2.5) +   #Week 04
#   annotate("text",x = 30, y = weeky, label = "W05 ", size = 2.5) +   #Week 05
   annotate("text",x = 37, y = weeky, label = "W06 ", size = 2.5) +   #Week 06
#   annotate("text",x = 44, y = weeky, label = "W07 ", size = 2.5) +   #Week 07
    annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
#   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
#   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
#   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
 #  annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
 #  annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
 #   annotate("text",x = 128, y = weeky, label = "W19 ", size = 2.5) +   #Week 19
   annotate("text",x = 135, y = weeky, label = "W20 ", size = 2.5) +   #Week 20
 #  annotate("text",x = 142, y = weeky, label = "W21 ", size = 2.5) +   #Week 21
   annotate("text",x = 149, y = weeky, label = "W22 ", size = 2.5) +   #Week 22
 #  annotate("text",x = 156, y = weeky, label = "W23 ", size = 2.5) +    #Week 23
  annotate("text",x = 163, y = weeky, label = "W24 ", size = 2.5) +    #Week 24
  
   annotate("text",x = 85, y = 50, label = "Feeding paused", size = 4, angle = 90, alpha = 0.6) +
    annotate("rect", ymin = 0, ymax = 100,  xmin = 82, xmax = 89,
             fill = "blue", alpha = .1) + 
  
 annotate("text",x = 64, y = 50, label = "Glycerol", size = 4, angle = 90, alpha = 0.6) +
    annotate("rect",ymin =0, ymax = 100, xmin = 62, xmax = 67, 
            fill = "blue", alpha = .1) +
  
    annotate("text",x = 94, y = 50, label = "Foam observed", size = 4, angle = 90, alpha = 0.65) +
    annotate("rect",ymin =0, ymax = 100, xmin = 93, xmax = 96, 
            fill = "blue", alpha = .1) +
  
  guides(fill=guide_legend(ncol=3)) + 
  ggtitle("Treatment")
```

### Combined
```{r fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
g1 <- ggarrange(p1, p2, common.legend = TRUE, widths = c(0.52,0.48), legend = "top")
g1 <- annotate_figure(g1, bottom = textGrob("Day of operation", gp = gpar(cex = 0.9)))

ggsave( "./Figures/Sup_areaplot_all_phylum.pdf", height=27, width=26, units='cm', dpi=300)
```

## Supp. figure S13 Ampviz heatmaps 16S top30
```{r message=FALSE, warning=FALSE}
psobject <- ps.flt
set.seed(1000)
psobject <- phyloseq::rarefy_even_depth(psobject)
psobject <- phyloseq::prune_samples(sample_data(psobject)$Date %in% c("2023-03-03","2023-03-15", "2023-04-12", "2023-04-26","2023-05-01",  "2023-05-15", "2023-05-24", "2023-06-02","2023-06-09","2023-06-05","2023-06-14","2023-06-28","2023-07-26","2023-08-09"), psobject)
psobject <- phyloseq::prune_samples(sample_data(psobject)$SampleID.TSVS %notin% c("F16", "F77", "F205", "F150"), psobject)
psobject <- phyloseq::prune_samples(sample_data(psobject)$DNATubeID %notin% c("#061", "#062","#063", "#064","#065","#066",
                                                                              "#216", "#217","#218", "#219","#220","#221"), psobject)

metadata <- data.frame(sample_data(psobject)) %>% 
  rownames_to_column("#OTU ID") 
taxtable <- data.frame(tax_table(psobject))
ASVtable <- data.frame(otu_table(psobject))
amp <- amp_load(otutable = ASVtable,
              metadata = metadata,
              taxonomy = taxtable)
p2 <- amp_heatmap(amp, tax_aggregate = "Phylum", tax_show = 30, showRemainingTaxa = TRUE, 
                  normalise = TRUE, facet_by = "SludgeType", group_by = "daysop")


p2 <- p2 +  theme(
  strip.text.x = element_text(size = 8, colour = "black", angle = 90),
  axis.text.x = element_text(size = 12)
  )

ggsave(plot = p2 , "./Figures/Sup_heatmap_v3v4midas53_all_phylum.png",  height=18, width=29, units='cm', dpi=200)
```



## Supp. figure S14 Ampviz heatmaps metagenome phyla
```{r message=FALSE, warning=FALSE}
library(ampvis2)

# Squeezemeta as described 
psobject <- psmg
sample_data(psobject)$SludgeType <- factor(sample_data(psobject)$SludgeType, levels = c("Control", "Treatment", "Foam", "Full-scale", "PSTWAS"))
#set.seed(1000)
#psobject <- phyloseq::rarefy_even_depth(psobject)
metadata <- data.frame(sample_data(psobject)) %>% 
  rownames_to_column("#OTU ID") 
#metadata$`#OTU ID` <-  str_replace_all(metadata$`#OTU ID`, "-", ".") 
taxtable <- data.frame(tax_table(psobject)) %>% dplyr::rename("Kingdom" = SuperKingdom)
ASVtable <- data.frame(otu_table(psobject))

amp <- amp_load(otutable = ASVtable,
              metadata = metadata,
              taxonomy = taxtable)
p <- amp_heatmap(amp, tax_aggregate = "Phylum", tax_show = 30, showRemainingTaxa = TRUE, 
                  normalise = TRUE, facet_by = "SludgeType", group_by = "daysop")
pmg <- p +  theme(strip.text.x = element_text(size = 8, colour = "black", angle = 90))

## HMMER Predicted, 16S extracted, Midas 53
df <- read.csv("./data/2024-06-11_arcbac_phylotabel.csv")
df <- df %>% 
    dplyr::rename(S561toS566_AD.AD6 = Chris.11,
      S1001_3_4_AD1_3_4 = Chris.12,
      S1002_5_6_AD2_5_6 = Chris.13,
      S1147_49_50_AD1_3_4 = Chris.14,
      S1148_51_52_AD2_5_6 = Chris.15,
      S1154_57_58_AD2_5_6 = Chris.16,
        F5 = Chris.01,
        F55 = Chris.02,
        F110 = Chris.03,
        F124 = Chris.04,
        F179 = Chris.05,
          S119.full = Chris.06,
          S370.full = Chris.07,
          S1273.ADFull = Chris.08,
          S1554.ADFull = Chris.09,
          S1813.ADFull = Chris.10) 

ncol(df %>% select( F5:S1154_57_58_AD2_5_6) )
colSums(df %>% select( F5:S1154_57_58_AD2_5_6) ) 
amp <- amp_load(otutable = df,
                metadata = (metadata) )

p <- amp_heatmap(amp, tax_aggregate = "Phylum", tax_show = 30, showRemainingTaxa = TRUE, 
                  normalise = TRUE, group_by = "daysop", facet_by = "SludgeType") #+ ggtitle("HMMER/sintax classification") 
phmmer.midas <- p +  theme(strip.text.x = element_text(size = 8, colour = "black", angle = 90))


g1 <- ggarrange(pmg, phmmer.midas, ncol = 1, align = "hv", labels = "AUTO")

ggsave(plot = g1, "./Figures/Sup_heatmap_hmmer_squemeta_phylum_comparison.png", height=33, width=25, units='cm', dpi=200)

```



## Supp. figure S15 Picrust2 heatmap
```{r message=FALSE, warning=FALSE}
library(circlize)
library(ComplexHeatmap)
output <- readRDS("./data/ancombcoutputpicrust-09Jul24")
res_prim = output$res
filtervec <- (res_prim %>% dplyr::filter(q_TreatmentTreatment <= 0.05))$taxon
ps.tmp <- ps_pw
ps.tmp <-  prune_taxa(taxa_sums(ps.tmp) > 10000, ps.tmp) # minimum reads per feature
ps.tmp <-  subset_samples(ps.tmp, sample_data(ps.tmp)$Period %in% c("Inhibition","Recovery/Feedingpause", "Recovery/Foaming"))
ps.tmp <-  subset_samples(ps.tmp, sample_data(ps.tmp)$SludgeType %in% c("Control", "Treatment"))
ps.tmp <-  prune_taxa(taxa_sums(ps.tmp) > 0, ps.tmp) # minimum reads per feature
ps.tmp <- microbiome::transform(ps.tmp, transform = "hellinger")
ps.tmp <- phyloseq::subset_taxa(ps.tmp, taxa_names(ps.tmp) %in% filtervec)

meta.tmp <- sample_data(ps.tmp)
sample_data(ps.tmp)$CSTR <- str_replace_all(sample_data(ps.tmp)$AD, c("AD1"="C1", 
                                       "AD2"="T1", 
                                          "AD3"="C2", 
                                            "AD4"="C3", 
                                              "AD5"="T2", 
                                               "AD6"="T3",
                                                "AD_full" = "Full-scale") )
vars <- c("CSTR", "Treatment")
sample_data(ps.tmp) <- tidyr::unite(data.frame(sample_data(ps.tmp)), "CSTR_Treat", all_of(vars),remove = FALSE )

summary(sample_data(ps.tmp)$Treatment)
summary(sample_data(ps.tmp)$AD)

# plot
taxtable <- data.frame(tax_table(ps.tmp))

colAnn <- HeatmapAnnotation(
        `Gasflow (ml/day)` = anno_barplot(sample_data(ps.tmp)$Gasflow_mL, width = unit(4, "cm")), 
        Digester = as.character(sample_data(ps.tmp)$Treatment),
        col = list(Digester= c( "Control" = "#4D98AC", "Treatment" = "#985C64")
                   ), annotation_name_gp= gpar(fontsize = 8), show_legend = FALSE
        )
set.seed(1232145)
row_ha <-  HeatmapAnnotation(
            which = c("row"),
            `MetaCyc class` = taxtable$CLASS1, annotation_name_gp= gpar(fontsize = 8),
            annotation_legend_param = list(`MetaCyc class` = list(ncol = 1)),
            show_legend = TRUE
              )
fig <- Heatmap((otu_table(ps.tmp)), 
        name = "Abundances", #title of legend
        row_names_gp = gpar(fontsize = 6.5),
        #split = sample_data(ps.tmp)$Treatment,
       #col_fun,
       show_heatmap_legend = FALSE,
        show_column_names = TRUE,
        show_row_names = TRUE,
       row_labels = c(data.frame(tax_table(ps.tmp))$Pathway2),
       column_labels = c(data.frame(sample_data(ps.tmp))$daysop),
       column_names_gp = gpar(fontsize = 6),
       cluster_columns = TRUE, 
       clustering_method_columns = "ward.D",
       top_annotation = colAnn,
       left_annotation = row_ha
    #row_km = 1, column_split = 2
       )    
       #column_split = 4

        # Text size for row names

png("./Figures/Sup_heatmap_metacyc_picrust_noleg.png", width = 19, height = 20,pointsize = 12,units="cm",res=300)
draw(fig, padding = unit(c(3, 3, 2, 7), "mm"), merge_legend = TRUE, heatmap_legend_side = c("top"),
     annotation_legend_side = "bottom") ## see right heatmap in following
```

## Supp. figure S16 - Metabolome PCA - is done in EcologyofFoaming_figure7

## Supp. figure S17 Picrust volcano plot
```{r message=FALSE, warning=FALSE}
colshm <- c("grey", "steelblue")
output <- readRDS("./data/ancombcoutputpicrust-09Jul24")
psAC <- ps_pw
psAC <-  subset_samples(psAC, sample_data(psAC)$Period %in% c("Inhibition","Recovery/Feedingpause", "Recovery/Foaming"))
psAC <-  subset_samples(psAC, sample_data(psAC)$SludgeType %in% c("Control", "Treatment"))
psAC <-  prune_taxa(taxa_sums(psAC) > 0, psAC)
res_prim = output$res
df = res_prim %>%
    dplyr::select(taxon, contains("Treatment")) %>% 
  dplyr::mutate('Label' = if_else(diff_TreatmentTreatment %in% c("TRUE"), taxon, NA_character_))
taxtable <- data.frame(tax_table(psAC)) %>% 
  rownames_to_column("taxon") 
df <- df %>% left_join(taxtable, by = "taxon")

# Overview
df %>% 
  group_by(diff_TreatmentTreatment) %>% 
  summarise(n())

p1 <- ggplot(data=df, aes(x=lfc_TreatmentTreatment, 
                          y=-log10(p_TreatmentTreatment), 
                          col = diff_TreatmentTreatment, 
                          label=taxon)) + 
  geom_point(alpha=0.75) + 
  scale_color_manual("Treatment effect", values = colshm) +
    geom_text_repel(max.overlaps = 10,show.legend = FALSE, size = 3) +
  ylab("-log10(p value)") +
  xlab("Log fold change") + 
  annotate(geom = 'text', label = paste("ANCOM with bias correction (n = 24 per group, paired for digester ID)\n  Pathway abundances (centred log-ratio) ~ Treatment"), 
             x = -Inf, y = -Inf, hjust = -0.05, vjust = -.2, size = 2.5) 

ggsave(plot = p1, "./Figures/Sup_volanoplot_picrust.png", height=7.75, width=13.5, units='cm', dpi=300)
rm(colshm)
```


## Supp. figure S18 Kegg pathways

### Aldex and create filtervec
```{r message=FALSE, warning=FALSE}
library(ALDEx2)
# Filter out full-scale AD
taxtable <- data.frame(tax_table(ps_kegg))
sample_data(ps_kegg)$SludgeType3 <- str_replace_all(sample_data(ps_kegg)$SludgeType, 
                                        c("Control"="RMITAD", 
                                       "Treatment"="RMITAD",
                                       "Foam"="RMITAD",
                                          "PSTWAS"="PSTWAS", 
                                            "Full-scale"="Full-scale") )
sample_data(ps_kegg)$SludgeType4 <- str_replace_all(sample_data(ps_kegg)$SludgeType, 
                                        c("Control"="Control", 
                                       "Treatment"="Treatment",
                                       "Foam"="Treatment",
                                          "PSTWAS"="PSTWAS", 
                                            "Full-scale"="Full-scale") )

ps.tmp   <- subset_samples(ps_kegg, sample_data(ps_kegg)$SludgeType3 %in% c("RMITAD"))
ps.tmp  = subset_taxa(ps.tmp, PathwayL1 %notin% c("Human Diseases", "Organismal Systems"))
ps.tmp   <-  prune_taxa(taxa_sums(ps.tmp) >= 100, ps.tmp) # minimum reads per feature
prev.kegg <- prevalencedf(ps.tmp , Pathway)

abun <- data.frame(phyloseq::otu_table(ps.tmp))
conds <- as.character(sample_data(ps.tmp)$SludgeType4)
summary(as.factor(conds))
x <- aldex.clr(abun, conds, mc.samples=1000, denom="all",
           verbose = TRUE)
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)
x.all <- data.frame(x.tt,x.effect) %>% rownames_to_column("ko")
png("./Figures/Sup_aldex2.png", width = 14, height = 12,pointsize = 12,units="cm",res=300)
aldex.plot(x.all, type="MA", test="wilcox")
dev.off()
#sgn <- sign(x.effect$effect.low) == sign(x.effect$effect.high)

#par(mfrow=c(1,2))
#plot(x.effect$rab.all, x.effect$diff.btw, pch=19, cex=0.3, col="grey", xlab="Abundance", ylab="Difference", main="Bland-Altman")
#points(x.effect$rab.all[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
#points(x.effect$rab.all[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")

#plot(x.effect$diff.win, x.effect$diff.btw, pch=19, cex=0.3, col="grey",xlab="Dispersion", ylab="Difference", main="Effect")
#points(x.effect$diff.win[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
#points(x.effect$diff.win[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")
#abline(0,2, lty=2, col="grey")
#abline(0,-2, lty=2, col="grey")

# filtervec
# by effect size
# filtervec <- (x.all %>% dplyr::filter(effect >= 2 |
#                          effect <= -2))

# by p value
filtervec <- (x.all %>% dplyr::filter(wi.eBH <= .05))
filtervec <- filtervec$ko
```

### Complex Heatmap
```{r}
library(circlize)
library(ComplexHeatmap)

prev.kegg <- prevalencedf(ps_kegg, Pathway)
sample_data(ps_kegg)$SludgeType3 <- str_replace_all(sample_data(ps_kegg)$SludgeType, 
                                        c("Control"="RMITAD", 
                                       "Treatment"="RMITAD",
                                       "Foam"="RMITAD",
                                          "PSTWAS"="PSTWAS", 
                                            "Full-scale"="Full-scale") )
sample_data(ps_kegg)$SludgeType3<- factor(sample_data(ps_kegg)$SludgeType3)
# filter
ps.tmp   <- subset_samples(ps_kegg, sample_data(ps_kegg)$SludgeType3 %in% c("RMITAD"))
ps.tmp <- microbiome::transform(ps.tmp, transform = "clr")
ps.tmp <- phyloseq::subset_taxa(ps.tmp, taxa_names(ps.tmp) %in% filtervec)
#ps.tmp <-  prune_taxa(taxa_sums(ps.tmp) > 0, ps.tmp) # minimum reads per feature

# plot
#col_fun = colorRamp2(c(min(otu_table(ps.tmp)), mean(otu_table(ps.tmp)), max(otu_table(ps.tmp))), c("black", "steelblue", "#FCFDBFFF"))
taxtable <- data.frame(tax_table(ps.tmp))
colAnn <- HeatmapAnnotation(
        #Date = as.character(sample_data(ps.tmp)$Date),
        `Day of operation` = as.character(sample_data(ps.tmp)$daysop),
        col = list(`Day of operation` = c( "61" = "#D2A277","75" = "#E495A5", 
                             "93" = "#72BB83"))
                        )
set.seed(1232145)
row_ha = rowAnnotation(`KEGG pathway` = taxtable$PathwayL2, 
                       annotation_name_gp= gpar(fontsize = 8))
#cols <- colorspace::rainbow_hcl(length(unique(sample_data(ps.tmp)$Date)))
#row_ano = rowAnnotation(foo = anno_block(gp = gpar(fill = 2:4)))

fig <- Heatmap((otu_table(ps.tmp)), 
        name = "Abundances", #title of legend
        row_names_gp = gpar(fontsize = 6.5),
        #split = sample_data(ps.tmp)$Treatment,
       #col_fun,
        show_column_names = TRUE,
        show_row_names = TRUE,
       row_labels = c(data.frame(tax_table(ps.tmp))$Pathway),
       column_labels = c(data.frame(sample_data(ps.tmp))$SludgeType),
       column_names_gp = gpar(fontsize = 6.5),
       clustering_method_columns = "ward.D",
       top_annotation = colAnn,
       left_annotation = row_ha,
    row_km = 1
       )    
       #column_split = 4
        # Text size for row names
fig
png("./Figures/Sup_heatmap_kegg_metagenome.png", width = 17, height = 12,pointsize = 12,units="cm",res=300)
draw(fig, padding = unit(c(3, 3, 2, 2), "mm"), merge_legend = TRUE) ## see right heatmap in following
dev.off()
# 550 x 850
```



## Supp. figure S19 violin/heatmap filaments 16S
### Violin
```{r message=FALSE, warning=FALSE}
ps.temp  <- ps.flt
set.seed(1234)
# summary(sample_sums(ps.temp))
# sort(sample_sums(ps.temp))
ps.temp <- phyloseq::rarefy_even_depth(ps.temp, sample.size = 30563, replace = FALSE)
#ps.temp <- ps.temp %>% microbiome::transform("clr")
#ps.temp <- microbiome::aggregate_taxa(ps.temp, level = "Genus") # faster
ps.temp <- ps.temp %>% microbiome::transform("compositional")
ps.summary <- ps.temp  %>% phyloseq::psmelt() %>% 
  dplyr::select(Date, SludgeType, week, Treatment, Abundance, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>% 
  dplyr::filter(Abundance > 0) %>% 
  group_by(Date,SludgeType, week, Treatment,  Kingdom, Phylum, Class, Order, Family,Genus, Species) %>% 
  summarise(Abundance = mean(Abundance) *100)  %>% unique()

ps.summary_sel <- ps.summary %>% 
  dplyr::filter( str_detect(Family, "Nocard") | str_detect(Family, "Microtrichaceae") | str_detect(Family, "Corynebacter") | 
                   str_detect(Family, "Mycobacter")  | str_detect(Family, "Tsukamurellaceae") | 
                                                                    str_detect(Genus, "Tetrasphaera") ) %>% 
                    dplyr::filter(week %in% c(10, 11,12, 13, 14,15,16))

p <- ggpubr::ggviolin(ps.summary_sel, x = "week", 
                    y = "Abundance",
                    ylab = "Abundance (%)",
                    xlab = "Week",
                   # facet.by = c("PMA_treat"),
                    fill = "SludgeType") +
                   # shape = "PMA_treat") + 
    theme_bw() +
  scale_fill_manual("Sample type",values = cols, labels = c("Control CSTR", "Treatment CSTR", "Foam", "Full-scale CSTR", "PS/TWAS")) +
    theme( panel.grid.major = element_blank(), panel.grid.minor = element_blank() ) +
   geom_point(aes(x =week,y=Abundance, fill=SludgeType), 
             position = position_jitterdodge(), alpha = 0.15)  +
   theme(axis.text.x = element_text(angle = 0, hjust = 0.5, size = 12))  +
  ggpubr::stat_compare_means(aes(x =week,y=Abundance, group=SludgeType,
                                 label = paste0(..p.signif..)),
                             method = "kruskal") 


```

### heatmap
```{r}
### heatmap of these 
ps.summary <- ps.temp  %>% phyloseq::psmelt() %>% 
  dplyr::select(OTU, Date, SludgeType, week, Treatment, Abundance, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>% 
  dplyr::filter(Abundance > 0) 
#ps.summary_sel <- ps.summary %>% 
#  dplyr::filter( str_detect(Family, "Nocard") | str_detect(Family, "Microtrichaceae") | str_detect(Family, "Corynebacter") | 
#                   str_detect(Family, "Mycobacter")  | str_detect(Family, "Tsukamurellaceae") ) %>% 
#                    dplyr::filter(Period %in% c("SteadyState", "Inhibition", "Recovery/Foaming", "Recovery/Feedingpause", 
#                                                "Glycerol"))
ps.summary_sel <- ps.summary %>% 
  dplyr::filter( str_detect(Family, "Nocard") | str_detect(Family, "Microtrichaceae") | str_detect(Family, "Corynebacter") | 
                   str_detect(Family, "Mycobacter")  | str_detect(Family, "Tsukamurellaceae") | 
                   str_detect(Genus, "Tetrasphaera") | str_detect(Genus, "Methanospirillum")) 

library(ampvis2)
ps.temp  <- ps.flt
psobject <- ps.temp
psobject <-   subset_samples(psobject, sample_data(psobject)$week %in% c(10, 11,12, 13,14,15,16))
psobject <- prune_taxa(taxa_sums(psobject) >  0, psobject) #minimum reads per feature
taxtable <- data.frame(tax_table(psobject)) 
ASVtable <- data.frame(otu_table(psobject))
metaobject <- data.frame(sample_data(psobject)) %>% 
  rownames_to_column("#OTU ID") 
amp <- amp_load(otutable = ASVtable,
              metadata = metaobject,
              taxonomy = taxtable)
filtervec <- ps.summary_sel$Genus %>% unique()
amp <- amp %>% amp_subset_taxa( normalise = TRUE, tax_vector = filtervec )
p1 <- amp_heatmap(amp, tax_aggregate = "Genus", tax_add = c("Family"),
                  group_by = "week", 
                  tax_show = 30, facet_by = "SludgeType",
                  showRemainingTaxa = TRUE, normalise = FALSE,
                  min_abundance = 0.001) +
  theme(axis.text.x = element_text(size = 12))

g <- ggarrange(p, p1, nrow = 2, labels = "AUTO")
g
ggsave(plot = g, "./Figures/Sup_violin_heatmap_filamen_hydrop.png", height=7, width=12, units='in', dpi=300)

```



## Supp. figure S20 heatmap filaments squeezemeta
```{r}
### ampviz heatmap 
ps.temp <- psmg
sample_sums(ps.temp)
sample_data(ps.temp)$SludgeType <- factor(sample_data(ps.temp)$SludgeType , levels = c("Control","Treatment","Foam","Full-scale", "PSTWAS"))
ps.summary <- ps.temp  %>% phyloseq::psmelt() %>% 
  dplyr::select(OTU, Date, SludgeType, week, Abundance, SuperKingdom, Phylum, Class, Order, Family, Genus, Species) %>% 
  dplyr::filter(Abundance > 0) 

#ps.summary_sel <- ps.summary %>% 
#  dplyr::filter( str_detect(Family, "Nocard") | str_detect(Family, "Microtrichaceae") | str_detect(Family, "Corynebacter") | 
#                   str_detect(Family, "Mycobacter")  | str_detect(Family, "Tsukamurellaceae") ) %>% 
#                    dplyr::filter(Period %in% c("SteadyState", "Inhibition", "Recovery/Foaming", "Recovery/Feedingpause", 
#                                                "Glycerol"))
ps.summary_sel <- ps.summary %>% 
  dplyr::filter( str_detect(Family, "Nocard") | str_detect(Family, "Microthrixaceae") | str_detect(Family, "Corynebacter") | 
                   str_detect(Family, "Mycobacter")  | str_detect(Family, "Tsukamurellaceae") | 
                   str_detect(Genus, "Tetrasphaera") | str_detect(Genus, "Methanospirillum")
                   )

library(ampvis2)
psobject <- ps.temp

#set.seed(12345)
#psobject <- phyloseq::rarefy_even_depth(psobject, replace = FALSE)
psobject <-   subset_samples(psobject, sample_data(psobject)$week %in% c(10, 11,12, 13,14,15,16))
psobject <- prune_taxa(taxa_sums(psobject) >  0, psobject) #minimum reads per feature
taxtable <- data.frame(tax_table(psobject)) 
ASVtable <- data.frame(otu_table(psobject))
metaobject <- data.frame(sample_data(psobject)) %>% 
  rownames_to_column("#OTU ID") 
amp <- amp_load(otutable = ASVtable,
              metadata = metaobject,
              taxonomy = taxtable)
filtervec <- ps.summary_sel$Genus %>% unique()
amp <- amp %>% amp_subset_taxa( normalise = TRUE, tax_vector = filtervec )
p1 <- amp_heatmap(amp, tax_aggregate = "Genus", tax_add = c("Family"),
                  group_by = "daysop", 
                  tax_show = 30, facet_by = "SludgeType",
                  showRemainingTaxa = TRUE, normalise = FALSE,
                  plot_values_size = 3,
                  min_abundance = 0.001) 
p1 <- p1 +  theme(strip.text.x = element_text(size = 8, colour = "black", angle = 90),
                  axis.text.x = element_text(size = 13))

ggsave(plot = p1, "./Figures/Sup_heatmap_filamen_hydrop.png", height=15, width=25, units='cm', dpi=300)
```


## Supp. figure S21 heatmap filaments kraken2
```{r}
### ampviz heatmap 
ps.temp <- pskr
sample_sums(ps.temp)
sample_data(ps.temp)$SludgeType <- factor(sample_data(ps.temp)$SludgeType , levels = c("Control","Treatment","Foam","Full-scale", "PSTWAS"))
ps.summary <- ps.temp  %>% phyloseq::psmelt() %>% 
  dplyr::select(OTU, Date, SludgeType, week, Abundance, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>% 
  dplyr::filter(Abundance > 0) 

#ps.summary_sel <- ps.summary %>% 
#  dplyr::filter( str_detect(Family, "Nocard") | str_detect(Family, "Microtrichaceae") | str_detect(Family, "Corynebacter") | 
#                   str_detect(Family, "Mycobacter")  | str_detect(Family, "Tsukamurellaceae") ) %>% 
#                    dplyr::filter(Period %in% c("SteadyState", "Inhibition", "Recovery/Foaming", "Recovery/Feedingpause", 
#                                                "Glycerol"))
ps.summary_sel <- ps.summary %>% 
  dplyr::filter( str_detect(Family, "Nocard") | str_detect(Family, "Microthrixaceae") | str_detect(Family, "Corynebacter") | 
                   str_detect(Family, "Mycobacter")  | str_detect(Family, "Tsukamurellaceae") )

library(ampvis2)
psobject <- ps.temp

#set.seed(12345)
#psobject <- phyloseq::rarefy_even_depth(psobject, replace = FALSE)
psobject <-   subset_samples(psobject, sample_data(psobject)$week %in% c(10, 11,12, 13,14,15,16))
psobject <- prune_taxa(taxa_sums(psobject) >  0, psobject) #minimum reads per feature
taxtable <- data.frame(tax_table(psobject)) 
ASVtable <- data.frame(otu_table(psobject))
metaobject <- data.frame(sample_data(psobject)) %>% 
  rownames_to_column("#OTU ID") 
amp <- amp_load(otutable = ASVtable,
              metadata = metaobject,
              taxonomy = taxtable)
filtervec <- ps.summary_sel$Genus %>% unique()
amp <- amp %>% amp_subset_taxa( normalise = TRUE, tax_vector = filtervec )
p1 <- amp_heatmap(amp, tax_aggregate = "Genus", tax_add = c("Family"),
                  group_by = "daysop", 
                  tax_show = 30, facet_by = "SludgeType",
                  showRemainingTaxa = TRUE, normalise = FALSE,
                  plot_values_size = 3,
                  min_abundance = 0.001) 
p1 <- p1 +  theme(strip.text.x = element_text(size = 8, colour = "black", angle = 90),
                  axis.text.x = element_text(size = 13))

ggsave(plot = p1, "./Figures/Sup_heatmap_filamen_hydrop_kraken.png", height=15, width=20, units='cm', dpi=300)
```


## Supp. figure S22 top 30 species Kraken2/Bracken
```{r}
psobject <- pskr
psobject <- phyloseq::subset_taxa(psobject, Kingdom %in% c("Archaea", "Bacteria"))
psobject <- prune_taxa(taxa_sums(psobject) > 0, psobject)
set.seed(1000)
psobject <- phyloseq::rarefy_even_depth(psobject)
metaobject <- data.frame(sample_data(psobject)) %>% 
  rownames_to_column("#OTU ID") 
#metadata$`#OTU ID` <-  str_replace_all(metadata$`#OTU ID`, "-", ".") 
taxtable <- data.frame(tax_table(psobject)) 
ASVtable <- data.frame(otu_table(psobject))

amp <- amp_load(otutable = ASVtable,
              metadata = metaobject,
              taxonomy = taxtable)

p1 <- amp_heatmap(amp, tax_aggregate = "Species", tax_add = c("Phylum"), tax_show = 30, showRemainingTaxa = TRUE, normalise = FALSE, group_by = "SludgeType2")
p2 <- amp_heatmap(amp, tax_aggregate = "Species", tax_add = c("Phylum"), tax_show = 30, showRemainingTaxa = TRUE, normalise = TRUE, 
                  group_by = "daysop", facet_by = "SludgeType") +
  theme(
    axis.text.x = element_text(size = 13)
  )
#g2 <- ggarrange(p1, p2)
p2
# sum(taxa_sums(psobject)) #15,219,740 reads
# 4,220,112 reads remaining
# sum(taxa_sums(psobject)) - 4220112 # 10999628 reads represented in figure
# 10999628 / 15219740 # 72.3% 

# sum(taxa_sums(psobject)) #9,930,576 23 July 24, including PSTWAS, full-scale, rarefied
# 1982212 remaining
# sum(taxa_sums(psobject)) - 1982212 #7948364 reads represented in figure
# 7948364 /  sum(taxa_sums(psobject)) #80%

ggsave(plot = p2, "./Figures/Sup_heatmap_kraken_RMITADs.png", height=15, width=31, units='cm', dpi=300)
```

## Supp. figure S23 top 30 species
```{r}
ps.onp <- ps.onp.ncbi.kraken
#summary(sample_sums(ps.onp))
set.seed(1000)
psobject <- phyloseq::rarefy_even_depth(ps.onp, replace = TRUE)
# ps.onp.ncbi (minimap) phyloseq-class experiment-level object
# otu_table()   OTU Table:         [ 4866 taxa and 8 samples ]
# sample_data() Sample Data:       [ 8 samples by 19 sample variables ]
# tax_table()   Taxonomy Table:    [ 4866 taxa by 8 taxonomic ranks ]

#ps.onp.ncbi.kraken - phyloseq-class experiment-level object
#otu_table()   OTU Table:         [ 5166 taxa and 8 samples ]
#sample_data() Sample Data:       [ 8 samples by 19 sample variables ]
#tax_table()   Taxonomy Table:    [ 5166 taxa by 8 taxonomic ranks ]

#ps.onp.st8 (kraken) - phyloseq-class experiment-level object
#otu_table()   OTU Table:         [ 2472 taxa and 8 samples ]
#sample_data() Sample Data:       [ 8 samples by 19 sample variables ]
#tax_table()   Taxonomy Table:    [ 2472 taxa by 8 taxonomic ranks ]

taxtable <- data.frame(tax_table(psobject)) 
ASVtable <- data.frame(otu_table(psobject))
metaobject <- data.frame(sample_data(psobject)) %>% 
  rownames_to_column("#OTU ID") 

metaobject$Sample_name <- str_replace_all(metaobject$Sample_name, c("CSTR1"="C1", 
                                       "CSTR2"="T1", 
                                          "CSTR3"="C2", 
                                            "CSTR4"="C3", 
                                              "CSTR5"="T2", 
                                               "CSTR6"="T3") )

amp <- amp_load(otutable = ASVtable,
              metadata = metaobject,
              taxonomy = taxtable) 
#p1 <- amp_heatmap(amp, tax_aggregate = "Genus", group_by = "SludgeType", tax_show = 30, showRemainingTaxa = TRUE, normalise = FALSE)

p <- amp_heatmap(amp, tax_aggregate = "Species", 
                  tax_add = c("Phylum"), 
                  facet_by = "SampleType",
                  group_by = "Sample_name", 
                  plot_values_size = 3,
                  tax_show = 50, showRemainingTaxa = TRUE, normalise = TRUE) 

#ggsave(polot = p, ggsave("./Figures/heatmap_16SFL.png", height=22, width=22, units='cm', dpi=300))
```

## Supp. figure  EC (not included) 
```{r}
firstdate <- 47  # selected periods for general manuscript
lastdate <- 124

weeky <- c(-0.5)
labely <- c(0.5)
segmenty <- c(0)

generate_vector <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 7
  }
  return(result)
}

generate_vectorday <- function(x, y) {
  result <- c()
  while (x <= y) {
    result <- c(result, x)
    x <- x + 1
  }
  return(result)
}


p <- foamdata %>% 
  dplyr::filter(SludgeType %notin% c("Full-scale")) %>%   # filter columns out
  ggline(x = "DayofOp",
         y = "EC", 
         color = "SludgeType",
         legend = "right",
         add = "mean_se",
         add.params = list(width = 1)) +
  # scale_color_manual(values = cols) +   # create a cols vector for custom colours
  theme_light() +
  scale_color_manual("Sludge Supernatent", values = cols)  +

    theme(
        axis.text.x = element_text(angle = 0,  hjust=0.5),
        legend.position = c(0.90, 0.825),
        legend.background=element_rect(fill = alpha("white", 0.5)),
        legend.ticks.length = element_blank(),
        legend.title = element_blank() )   + 
  
  scale_x_continuous(breaks = generate_vector(firstdate,lastdate),
                     minor_breaks = generate_vectorday(firstdate,lastdate),
               limits = c(firstdate, lastdate) )  +

  annotate("text",x = 51, y = weeky, label = "W08 ", size = 2.5) +   #Week 08
   annotate("text",x = 58, y = weeky, label = "W09 ", size = 2.5) +   #Week 09
   annotate("text",x = 65, y = weeky, label = "W10 ", size = 2.5) +   #Week 10
   annotate("text",x = 72, y = weeky, label = "W11 ", size = 2.5) +   #Week 11
   annotate("text",x = 79, y = weeky, label = "W12 ", size = 2.5) +   #Week 12
   annotate("text",x = 86, y = weeky, label = "W13 ", size = 2.5) +   #Week 13
   annotate("text",x = 93, y = weeky, label = "W14 ", size = 2.5) +   #Week 14
   annotate("text",x = 100, y = weeky, label = "W15 ", size = 2.5) +   #Week 15
   annotate("text",x = 107, y = weeky, label = "W16 ", size = 2.5) +   #Week 16
   annotate("text",x = 114, y = weeky, label = "W17 ", size = 2.5) +   #Week 17
   annotate("text",x = 121, y = weeky, label = "W18 ", size = 2.5) +    #Week 18
  
  
  annotate("text",x = 54, y = labely, label = "Steady state", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 47, xend = 62,
            arrow = arrow(ends = "last", angle = 90, length = unit(.1,"cm"))) +
  
 #  annotate("text",x = 65, y = labely, label = "Glycerol   ", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 62, xend = 67,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 74, y = labely, label = "   Inhibition", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 67, xend = 82,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  annotate("text",x = 86, y = labely, label = "Recovery", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 82, xend = 90,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
  #annotate("text",x = 94, y = labely, label = "Foaming ", size = 2.5) +
  annotate("segment", y = segmenty, yend = segmenty, x = 90, xend = 98,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 101, y = labely, label = "    Recovery", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 98, xend = 106,
            arrow = arrow(ends = "both", angle = 90, length = unit(.1,"cm"))) +
  
    annotate("text",x = 115, y = labely, label = "Steady state post foam", size = 2.5) +
   annotate("segment", y = segmenty, yend = segmenty, x = 98, xend = 124,
            arrow = arrow(ends = "first", angle = 90, length = unit(.1,"cm"))) +
  
   annotate("text",x = 85, y = 4, label = "Feeding paused", size = 4, angle = 90, alpha = 0.5) +
    annotate("rect", ymin = 0, ymax = 8, xmin = 82, xmax = 89, fill = "blue", alpha = .1) +

      annotate("text",x = 64, y = 4, label = "Glycerol", size = 4, angle = 90, , alpha = 0.5) +
    annotate("rect",ymin =0, ymax = 8, xmin = 62, xmax = 67, 
             fill = "blue", alpha = .1) +

    annotate("text",x = 93.5, y = 4, label = "Foam observed", size = 3.5, angle = 90, , alpha = 0.5) +
    annotate("rect", ymin = 0, ymax = 8, xmin = 91, xmax = 96, 
             fill = "blue", alpha = .1) +
  
      ylab("EC (S/cm)") +
      xlab("Day of operation")

ggsave("./Figures/Sup_lineplot_ec_daysop.png", height=2.5, width=6, units='in', dpi=300)
```

